---
title: Hands-on Machine Learning with R
geometry: "left=3.5cm,right=3.5cm,top=2cm,bottom=2cm"
output:
  pdf_document: 
  number_sections: true
---
\setcounter{section}{3}
# Linear Regression

```{r, global_options, include=FALSE}
library(knitr)
opts_chunk$set(warning=FALSE, message=FALSE)
```

*Linear  regression* is one of the simplest algs for supervised learning. But it's a good starting point and many more complex methods can be seen as extensions of it. 

## Prerequisites

Adding `vip` packages for interpretability of variable importance. Ames data set from before. 
```{r, echo = FALSE}
# Helper packages
library(dplyr)    # for data manipulation
library(ggplot2)  # for awesome graphics

# Modeling packages
library(rsample)  # to split the set
library(caret)    # for cross-validation, etc.

# Model interpretability packages
library(vip)      # variable importance

# viz packages
library(scales) # for transparency
library(rsm) # for contour plots

# load ames housing data
ames <- AmesHousing::make_ames()

# Stratified sampling with the rsample package
set.seed(123)
split <- initial_split(ames, prop = 0.7, 
                       strata = "Sale_Price")
ames_train  <- training(split)
ames_test   <- testing(split)
```


## Simple Linear Regresison

SLR assumes the relationship between two continuous variables is at least approximately linear. 

$$Y_i=\beta_0+\beta_1 X_i+\epsilon_i, \qquad for \quad i=1,2,...,n,$$
Where $Y_i$ represents the response/target variable, $X_i$ is the $i^{th}$ feature value and teh betas are fixed but unknown constants (coefficients or parameters), representing the intercept and the slope.

The $\epsilon_i$ term represents noise or random error. Here we assume the errors have a mean of zero and constant variance $\sigma^2$. This is denoted as $\stackrel{iid}{\sim}N(0, \sigma^2)$. Since the errors are centered on zero - the expected value $E(\epsilon_i) = 0$, linear regression is really a problem of estimating the conditional mean:

$$E(Y_i|X_i) = \beta_0+\beta_1 X_i$$
WHich we can shorten to just $E(Y)$. So the interpretation is in therms of *average responses*. E.g. $\beta_0$ is the average response value when $X=0$ - sometimes referred to as the *bias term* and $\beta_1$ is the increase in the average response if $X$ increases by one unit, aka the *rate of change*. 

### Estimation

We want the best fitting line, but what is the best fit? The most common way, called *Oridnary least squares* (OLS) is to minimise the *residual sum of squares*:

$$RSS(\beta_0,\beta_1)=\sum^{n}_{i=1}[Y_i-(\beta_0+\beta_1X_i)]^2=\sum^{n}_{i=1}(Y_i-\beta_0-\beta_1X_i)^2.$$
We denote the OLS stimates of the coefficients as $\hat \beta_0$ and $\hat \beta_1$. Once we have the estimated regression equation, we can predict values of $Y$ for $X_{new}$:

$$\hat Y_{new} = \hat \beta_0 + \hat \beta_1 X_{new}$$
Where $\hat Y_{new}$ is the estimated mean response at $X = X_{new}$. 

So let's try modelling the ames data relationship between the sale price and the above ground living area. [This link)](https://drsimonj.svbtle.com/visualising-residuals)has good info on visualising residuals. 

```{r}
model1 <- lm(Sale_Price ~Gr_Liv_Area, data = ames_train)

# let's have a look at the resisiduals and plot them 

x <- ames_train
x$predicted <- predict(model1)   # Save the predicted values
x$residuals <- residuals(model1) 

# this is what the data looks like
x %>% select(Sale_Price, predicted, residuals) %>% head()

# here are some plots
par(mfrow = c(1,2))
par(mar = c(4,4,2,0.1)+0.10)
plot(x$Gr_Liv_Area, x$Sale_Price, col = alpha("black", 0.3), pch = 20,
     main = "fitted regression line")
abline(model1, col = "cadetblue4", lwd = 2)
plot(x$Gr_Liv_Area, x$Sale_Price, col = alpha("black", 0.3), pch = 20,
     main = "fitted regression line with residuals")
abline(model1, col = "cadetblue4", lwd = 2)
for (i in 1: nrow(x)){
lines(c(x$Gr_Liv_Area[i], x$Gr_Liv_Area[i]),
      c(x$Sale_Price[i], x$predicted[i]), 
      lwd = 0.5)}
```


Use `coef()` and `summary()` to have a look a the coefficients. !!! I don't get the same data, even though i have the same seed in the split?

```{r}
coef(model1)
summary(model1)
# glimpse(model1)
```

So we estimate that an increase in area by one square foot increases the selling price by 114.88$. SO nice and intuitive. 

One drawback  of using least squares is that we only have estimates of the coefficients, but not of the error variance $\sigma^2$. LS makes no assumptions about the random errors, so we cannot estimate $\sigma^2$. 

An alternative is to use *maximum likelihood* estimation (ML) to estimate $\sigma^2$ -- which we need to characterise the variability of our model. For ML we have to assume a particular distribution of the errors, most commonly that they are normally distributed. Under these assumptions the estimate of the error variance is 

$$\hat\sigma^2=\frac{1}{n-p} \sum^n_{i=1}(Y_i-\hat Y_i)^2 = \frac{1}{n-p} \sum^n_{i=1}r_i^2$$

Where  $r_i$ is the residual of the $i^{th}$ observation. and $p$ is the number of parameters or coefficients in the model. \hat\sigma^2 is also known as the mean squared error (MSE) and it's square root is the RMSE, and you can get it out of an `lm` object using `sigma()`

```{r}
sigma(model1)
sigma(model1)^2
```
!!! the sigma is slightly different from the RMSE reported in the summary, not sure why, same in book. 

### Inference

The coefficients are only point estimates, so that's not super useful without a measure of variability. This is usually measured with a *standard error* (SE), the square root of it's variance. If we assume the errors are distributed  $\stackrel{iid}{\sim}N(0, \sigma^2)$ ,then the SEs for the coefficients are simple and are expressied under the `Std. Error` heading in the summary for the model.

From the SE we can also do a t-test to see if the coefficients are statistically significantly different from zero. (!!! stantistically significant from zero is probably wrong). 

The t-statistic is simply the estimated coefficient divided by the SE, which measures the number of standard deviations each coefficient is away from zero. The p-values are reported in the same table. 

Under these same assumptions we can also derive the confidence intervals for the $\beta$ coefficients. The formual is:

$$\beta_j \pm t_{1-α/2,n-p} \hat {SE}(\hat \beta_j)$$

In R you can construct them using `confitn()`

```{r}
confint(model1, level = 0.95)

# or if you wanna spell it out:
coef(model1)[2] - qt(0.975, model1$df.residual)*coef(summary(model1))[2,2]
coef(model1)[2] + qt(0.975, model1$df.residual)*coef(summary(model1))[2,2]
```

So with 95% confidence we estimate that the mean sale price goes up between 109 and 119$ for each additional square foot. 

Don't forget that these SEs and t-stats etc in the summary are based on the following assumptions:

1. Independent observations
2. The random errors have mean zero, and constant variance
3. The random errors are normally distributed

If your data deviate from these assumptions, there are some remedial actions you can take.. 

## Multiple linear regression

Extend the simple linear regression with more predictors to see e.g. how are and year built are (linearly) related to the sales price using *mulitple linear regression* (MLR). 

$$Y_i=\beta_0+\beta_1 X_i+\beta_2 X_2 + \epsilon_i, \qquad for \quad i=1,2,...,n,$$

Which you do in R by using `+` to separate predictors:



```{r}
(model2 <- lm(Sale_Price ~ Gr_Liv_Area + Year_Built, data = ames_train))

# or use update
(model2 <- update(model1, .~. + Year_Built))

```

So holding the year constant, each additional square foot of living area increses the mean selling price by 99\$. And holding the area constant, each additional year the home is newer by increases the mean price by 1093\$. Here are some contour plots:


```{r, fig.height = 3, echo = FALSE, fig.cap = "Contour plot of the fitted regression surface"}
contour(model2,  Year_Built ~ Gr_Liv_Area ,
        image = TRUE, img.col = rev(heat.colors(40)),
        main = "contour plot for model 2")

model3 <- lm(Sale_Price ~ Gr_Liv_Area + Year_Built + Gr_Liv_Area : Year_Built, data = ames_train)
contour(model3,  Year_Built ~ Gr_Liv_Area ,
        image = TRUE, img.col = rev(heat.colors(40)),
        main = "contour plot for model 3")

```

The left one only has main effects and is therefore flat. Including interaction effects models curvature: the effect of one predictor now depend on the level of the other. So in our example this would mean including the product of both predictors:


$$Y_i=\beta_0+\beta_1 X_i+\beta_2 X_2 + \beta_3 X_1 X_2 \epsilon_i, \qquad for \quad i=1,2,...,n,$$

In R the formula is either `y ~ x1 + x2 ~ x1:x2` or `y ~ x1 * x2`. 

Note the *hierarchy principle* which means that any lower order terms corresponding to the interaction term must also be included in the model. 

You can include as many predictors as you like - as long as you have more observations than predictors! (So in wide tables you cannot include all of them!). These can also be interactions, or transoformations: e.g. $X_3 = X_1X2$ or $X_4 = \sqrt(X_3)$. Of course after two  dimensions visualisation becomes impractical, because we have a hyperplane of best fit.

We can try all of the predictors in the data set and clean up the output using the `broom` package: (!!! again, the results are even more different than before). 

```{r}
model4 <- lm(Sale_Price ~ ., data = ames_train)

broom::tidy(model4)
```

## Assessing model accuracy

